On the Gravitational stability of gravito-turbulent accretion disks

### 2016-01-11 paper/model3d.tex
regular update

We adopt the viscosity law  
\begin{align}
  \nu = \alpha
  \frac{c_{s0}^2}{\Omega}\left(\frac{\rho}{\rho_\mathrm{eq}}\right)^\mu\left(\frac{P}{P_\mathrm{eq}}\right)^\lambda,            
\end{align}
where subscript `eq' denotes the equilibrium state and  
%$c_{s0}$ is a modified sound-speed defined below. 
$c_{s0}^2\equiv P_\mathrm{eq}(0)/\rho_\mathrm{eq}(0)$. 
The dimensionless coefficient
$\alpha=\alpha(\rho_\mathrm{eq},P_\mathrm{eq})$ characterizes the
magnitude of the shear viscosity \emph{in steady state} 
and is time-independent. The indices $\mu,\,\lambda$ are free
parameters chosen to model how the viscosity behaves in the perturbed
state.  
%Given the uncertain
%properties of gravito-turbulence, we will treat $\mu,\lambda$ as free
%parameters. 
We adopt the same prescription for the bulk viscosity but
with $\alpha\to\alpha_b$.

Our numerical calculations use $\mu=-1,\lambda=0$ so 
that $\rho\nu$ remains constant, as adopted in previous, pioneering
studies of viscous GI 
\citep{lynden-bell74,hunter83,willerding92,gammie96}.  
This choice also conveniently eliminates viscous over-stability
\citep{schmit95,latter06}, which is unrelated to self-gravity, that
would otherwise contaminate our results.

%the alpha(rho_eq, P_eq) function can be derived with some rigor 
%but this assumes steady state. don't know how to model
%time-dependence.

While steady-state viscosity values can be determined analytically
\citep[e.g.][]{rafikov15}, its time-dependent behavior is not
well-explored. \cite{laughlin96b} suggested a $\nu \propto 1/\Sigma$
dependence when modeling the evoluting of self-gravitating disks 
as a viscous process. \cite{paardekooper12} measured
an increase in the effective viscosity when their gravito-turbulent
disk models underwent fragmentation (i.e. as $\Sigma$ increases).  
These motivate the choice $\mu=-1,\lambda=0$, but we emphasize this is
done to bring out the physical process of interest --- viscosity-assisted
gravitational instability.

@@

\subsection{Steady states and cooling models}

We consider equilibrium solutions (omitting the `eq' subscripts
hereafter for clarity)
\begin{align}
  \bm{v} &= -q\Omega x \hat{\bm{y}}, \\
  \rho   &= \rho(z),\\
  P      &= P(z) \equiv c_s^2(z)\rho.
\end{align} 
%and assume a polytropic relation for the equilibrium pressure,
%\begin{align}
%  P  =
%  \frac{c_{s0}^2\rho_0}{\Gamma}\left(\frac{\rho}{\rho_0}\right)^\Gamma
%  \equiv \frac{c_s^2(z)}{\Gamma}\rho,
%\end{align}
%where $\rho_0$ is the equilibrium midplane density and $\Gamma$ is the
%constant polytropic index.  
The equilibrium density and pressure fields are obtained by solving
the vertical momentum equation with self-gravity,
\begin{align}
  &\frac{1}{\rho}\frac{dP}{dz} +
  \Omega_z^2z + \frac{d\Phi}{dz} = 0, \label{vert_eq1}\\
 &\frac{d^2\Phi}{dz^2} = 4 \pi G \rho,\label{vert_eq2}
\end{align}
%A numerical solution is usually required to obtain the equilibrium
%state, except for special values of $\Gamma$. 
together with thermodynamic equilibrium,
%\subsection{Thermal equilibrium}
%In equilibrium the energy equation implies
\begin{align}\label{thermal_eq}
%  \mathcal{H}_\mathrm{visc} + \mathcal{H}_\mathrm{ext} = \Lambda.   
(q\Omega)^2\rho\nu + \mathcal{H}_\mathrm{ext} = \Lambda,
\end{align}
where the first term represents viscous heating. For the viscous
problem we may set $\mathcal{H}_\mathrm{ext}=0$. However, if we wish
to neglect viscosity but include cooling, we must invoke
$\mathcal{H}_\mathrm{ext}\neq0$ to define an equilibrium state.  
To proceed further, we separately describe the two cooling models
considered in this work. 
%It is possible to have both models present - but too complicated

\subsubsection{Beta cooling}
In this case the energy loss per unit volume is specified 
as an explicit function of the thermodynamic variables. 
A prototypical example is 
\begin{align}\label{beta_cool}
  \Lambda =
  %  \frac{P}{(\gamma-1)T}\frac{\left(T-T_\mathrm{irr}\right)}{t_c}, 
  \frac{\mathcal{R}\rho}{(\gamma-1)}\frac{\left(T-T_\mathrm{irr}\right)}{t_c}, 
\end{align}
where 
\begin{align}
  \tcool = \beta\Omega^{-1} 
\end{align}
is the cooling timescale with $\beta$ a constant input parameter, and
\begin{align}
  \tirr = \theta T_\mathrm{eq}
\end{align}
is a reference temperature field parameterized as a constant fraction
$\theta$ of the equilibrium temperature $T_\mathrm{eq}$. We will refer
to $\theta$ as the `irradiation level'.

This type of cooling is widely applied in 2D and 3D
numerical simulations of self-gravitating disks \citep{gammie01,
  rice05,baehr15}. It permits numerical experiments to be carried out in a
controlled manner as a function of the cooling time $\beta$.

We remark that for the linear problem, 
any explicit cooling function $\Lambda=\Lambda(\rho,T)$ can be
mapped to that obtained from the standard function in Eq.
\ref{beta_cool}. We do this in \S\ref{2dppd} where we adopt a more
realistic beta cooling prescription for 2D PPDs.

For models with beta cooling we assume an equilibrium polytropic relation
\begin{align} 
  P  =
%  \frac{c_{s0}^2\rho_0}{\Gamma}\left(\frac{\rho}{\rho_0}\right)^\Gamma,
  %  \equiv \frac{c_s^2(z)}{\Gamma}\rho,
c_{s0}^2\rho_0\left(\frac{\rho}{\rho_0}\right)^\Gamma,
\end{align}
where $\Gamma$ is the constant polytropic index and $\rho_0 =
\rho(0)$. The vertical
structure is first obtained from Eq. \ref{vert_eq1}---\ref{vert_eq2}
(usually numerically), then inserted into
Eq. \ref{thermal_eq} to infer the required viscosity profile

@@

@@

While radiative cooling is arguably more realistic than 
parametrized beta cooling, it complicates our 
problem because it generally implies a vertically non-isothermal
equilibrium disk, and increases the order of the linearized equations.

@@

%\subsection{Steady states}
%\subsubsection{Equilibria with beta cooling}   
%\subsubsection{Equilibria with radiative cooling}
%If $\mathcal{H}_\mathrm{ext}=0$ then viscosity and cooling
%cannot be prescribed independently and we have 
%\begin{align}\label{alpha_beta_relation}
%  \alpha(z) = \frac{1}{\Gamma(\gamma-1)\beta
%    q^2}\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left(1 - \theta\right). 
%\end{align}
%On the other hand, if a time-independent heat source/sink is invoked
%($\mathcal{H}_\mathrm{ext}\neq 0$), then we may choose $\alpha, \beta$
%and $\theta$ separately.

%\subsection{Parameters}
%{\bf temporary, maybe move to intro}
%For the 3D problem it is convenient to parameterize the strength of
%self-gravity via 
%\begin{align}
%  Q_\mathrm{3D} \equiv \frac{\Omega^2}{4\pi G \rho_0}. 
%\end{align}
%It is also useful to define the
%two-dimensional (2D) Toomre parameter 
%\begin{align}
%  Q = \frac{c_{s0}\Omega}{\pi G \Sigma}, 
%\end{align}
%where $\Sigma = \int_{-\infty}^\infty\rho dz$ is the surface density.

%We are interested in comparing systems with different $\beta$. 
%However, this gives unphysically large values of $\alpha$ for small
%values of $\beta$. In order to compare systems with different $\beta$,
%we adopt the following procedure. We choose parameters  
%$\theta_\mathrm{min}$ and $\alpha_\mathrm{max}$, and evaluate
%Eq. \ref{alpha_beta_relation} at the midplane to define a cooling time
%$\beta_\mathrm{min}$. Then for $\beta\geq\beta_\mathrm{min}$ we use 
%Eq. \ref{alpha_beta_relation} to set $\alpha$ at fixed
%$\theta=\theta_\mathrm{min}$; while for $\beta < \beta_\mathrm{min}$
%we use Eq. \ref{alpha_beta_relation} to set $\theta$ at fixed
%$\alpha(0)=\alpha_\mathrm{max}$.  That is, we raise the reference
%temperature (i.e. irradiation levels) when the imposed cooling is too
%rapid to be sustaind by viscous heating.

>>>>

We adopt the viscosity law  
\begin{align}
  \nu = \alpha
  \frac{c_{s0}^2}{\Omega}\left(\frac{\rho}{\rho_\mathrm{eq}}\right)^\mu\left(\frac{P}{P_\mathrm{eq}}\right)^\lambda,            
\end{align}
where subscript `eq' denotes the equilibrium state and  
%$c_{s0}$ is a modified sound-speed defined below. 
$c_{s0}^2\equiv P_\mathrm{eq}(0)/\rho_\mathrm{eq}(0)$. 
The dimensionless coefficient
$\alpha=~\alpha(\rho_\mathrm{eq},P_\mathrm{eq})$ characterizes the
magnitude of the shear viscosity \emph{in steady state}. The indices
$\mu,\,\lambda$ are free parameters chosen to model how the viscosity
behaves in the perturbed state.  
%Given the uncertain
%properties of gravito-turbulence, we will treat $\mu,\lambda$ as free
%parameters. 
We adopt the same prescription for the bulk viscosity but
with $\alpha\to\alpha_b$.

Our numerical calculations use $\mu=-1,\lambda=0$ so 
that $\rho\nu$ is time-indpendent, as adopted in previous 
studies of viscous GI 
\citep{lynden-bell74,hunter83,willerding92,gammie96}.  
This choice also conveniently eliminates viscous over-stability
\citep{schmit95,latter06}, which is unrelated to self-gravity, that
would otherwise contaminate our results. %obscure

While steady-state viscosity values can be determined analytically
\citep[e.g.][]{rafikov15}, its time-dependent behavior is not
well-explored. We emphasize the choice $\mu=-1,\lambda=0$ is made to
bring out the physically relevant process --- viscosity-assisted
gravitational instability. %expect to be generally present, but will
                           %be obscured by viscous over-stability with
                           %other indices
Interestingly, though, \cite{laughlin96b} have suggested a $\nu \propto 1/\Sigma$
dependence when modeling the evoluting of self-gravitating disks 
as a viscous process. Furthermore, \cite{paardekooper12} measured
an increase in the effective viscosity when their gravito-turbulent
disk models underwent fragmentation (i.e. as $\Sigma$ increases).

@@

\subsection{Steady states and cooling models}
We consider equilibrium solutions (here omitting the `eq' subscripts
for clarity) 
\begin{align}
  \bm{v} &= -q\Omega x \hat{\bm{y}}, \\
  \rho   &= \rho(z),\\
  P      &= P(z) \equiv c_s^2(z)\rho.
\end{align} 
%and assume a polytropic relation for the equilibrium pressure,
%\begin{align}
%  P  =
%  \frac{c_{s0}^2\rho_0}{\Gamma}\left(\frac{\rho}{\rho_0}\right)^\Gamma
%  \equiv \frac{c_s^2(z)}{\Gamma}\rho,
%\end{align}
%where $\rho_0$ is the equilibrium midplane density and $\Gamma$ is the
%constant polytropic index.  
The equilibrium density and pressure fields are obtained by solving
the vertical momentum equation with self-gravity,
\begin{align}
  &\frac{1}{\rho}\frac{dP}{dz} +
  \Omega_z^2z + \frac{d\Phi}{dz} = 0, \label{vert_eq1}\\
 &\frac{d^2\Phi}{dz^2} = 4 \pi G \rho,\label{vert_eq2}
\end{align}
%A numerical solution is usually required to obtain the equilibrium
%state, except for special values of $\Gamma$. 
together with thermodynamic equilibrium,
%\subsection{Thermal equilibrium}
%In equilibrium the energy equation implies
\begin{align}\label{thermal_eq}
%  \mathcal{H}_\mathrm{visc} + \mathcal{H}_\mathrm{ext} = \Lambda.   
(q\Omega)^2\rho\nu + \mathcal{H}_\mathrm{ext} = \Lambda,
\end{align}
where the first term represents viscous heating. For the viscous
problem, $\nu\neq0$, we may set $\mathcal{H}_\mathrm{ext}=0$. However,
if we wish to neglect viscosity but include cooling, we must invoke
$\mathcal{H}_\mathrm{ext}\neq0$ to define an equilibrium state.  
To proceed further, we separately describe the two cooling models
considered in this work. 
%It is possible to have both models present - but too complicated

\subsubsection{Beta cooling}
In this case the energy loss per unit volume is specified 
as an explicit function of the thermodynamic variables. 
A prototypical example is 
\begin{align}\label{beta_cool}
  \Lambda =
  %  \frac{P}{(\gamma-1)T}\frac{\left(T-T_\mathrm{irr}\right)}{t_c}, 
  \frac{\mathcal{R}\rho}{(\gamma-1)}\frac{\left(T-T_\mathrm{irr}\right)}{t_c}, 
\end{align}
where 
\begin{align}
  \tcool = \beta\Omega^{-1} 
\end{align}
is the cooling timescale with $\beta$ a constant input parameter, and
\begin{align}
  \tirr = \theta T_\mathrm{eq}
\end{align}
is a reference temperature field parameterized as a constant fraction
$\theta$ of the equilibrium temperature $T_\mathrm{eq}$. We will refer
to $\theta$ as the `irradiation level'.  
%This cooling function restores 
%the disk temperature to $\tirr$ over a time-scale $\tcool$.

   Beta cooling is widely applied in 2D and 3D
numerical simulations of self-gravitating disks \citep{gammie01,
  rice05,baehr15}. It permits numerical experiments to be carried out in a
controlled manner as a function of the cooling time $\beta$.

For models with beta cooling we assume an equilibrium polytropic relation
\begin{align} 
  P  =
%  \frac{c_{s0}^2\rho_0}{\Gamma}\left(\frac{\rho}{\rho_0}\right)^\Gamma,
  %  \equiv \frac{c_s^2(z)}{\Gamma}\rho,
c_{s0}^2\rho_0\left(\frac{\rho}{\rho_0}\right)^\Gamma,
\end{align}
where $\Gamma$ is the constant polytropic index and $\rho_0 =
\rho(0)$. The vertical
structure is first obtained from Eq. \ref{vert_eq1}---\ref{vert_eq2}
then inserted into Eq. \ref{thermal_eq} to infer the required
viscosity profile,

@@

We remark that for the linear problem  
any explicit cooling function $\Lambda=\Lambda(\rho,T)$ can be
mapped to that obtained from the standard function in Eq.
\ref{beta_cool}. We do this in \S\ref{2dppd} where we adopt a more
realistic beta cooling prescription for 2D PPDs.

@@

@@

While radiative cooling is arguably more realistic than
parametrized beta cooling, it generally implies a vertically
non-isothermal equilibrium disk, and increases the order of the
linearized equations. It also only applies to optically-thick 
disks. It is possible to modify the flux function to account for
optically-thin disks, but this complication is beyond the scope of
this work.

### 2016-01-11 paper/results2d.tex
regular update

\section{Two-dimensional disks with beta cooling}
We begin in the 2D limit with beta cooling  to make connection 
with previous studies. The disk
material assumed to be confined to the midplane so $\delta v_z=0$. 
We make the replacements 
$\rho \to \Sigma$ and $P\to\Pi$, the vertically-integrated
pressure, and set $\Gamma=1$ without loss of generality. 
%The perturbations $\delta\Sigma$, 
%$\delta\Pi$, $\delta v_x$ and $\delta v_y$ have no vertical
%dependence and $\delta v_z=0$. 
The gravitational potential perturbation remains 3D and its midplane
value is given by    
\begin{align}
  \delta \Phi(z=0) = -\frac{2 \pi G}{|k|}\delta\Sigma
\end{align}
\citep{shu70}.

@@

\subsection{Inviscid limit}\label{2d_inviscid}
Here we neglect viscosity by setting 
$\alpha = \alpha_b = 0$. This implies an external,
time-independent heat source must be invoked to balance the imposed
cooling rate in equilibrium ($\mathcal{H}_\mathrm{ext}\neq 0$). Alternatively, we are assuming 
viscosity is only important for setting up thermal equilibrium, and
can be neglected for the pertubations. This is useful as the resulting
instabilties are only due to self-gravity.  In any case, we are
considering the \emph{sole} effect of cooling on stability.

@@

Eq. \ref{inviscid} is a cubic equation in $S$. Applying the
Routh-Hurwitz criterion, we find that stability is ensured if 
\begin{align}
  \gamma > \theta \quad \text{and} \quad 
  Q > \frac{1}{\sqrt{2\theta(2-q)}} 
\end{align}
are both satisfied. Thus, the question of stability does not depend on
the cooling rate $\beta$. The system becomes more unstable with
decreasing $\theta$.

@@

\subsubsection{$\theta = 0$ and $Q=Q_\mathrm{crit}$}  
Beta cooling with $\theta=0$ is typically employed in numerical
simulations of self-gravitating disks \citep{gammie01}. There, 
a steady state corresponds to the imposed cooling entirely balanced by heat generated through 
gravito-turbulence. In the present framework, this gravito-turbulence
is assumed to provide a constant 
background heating ($\mathcal{H}_\mathrm{ext}$) and does
not enter the linearized equations. Cooling, however, is still imposed in the perturbed state. 
From the discussion above, this system is unconditionally unstable for
finite $Q$. The most unstable wavelength decreases with decreasing
$\beta$. The $\beta\to0$ limit corresponds to a pressureless disk.

@@

Specifically, let us consider a disk with 
\begin{align}\label{qcrit_def}
  \qtwo = \frac{1}{\sqrt{2\gamma(2-q)}}\equiv Q_{\mathrm{crit}},
\end{align} 
so that it is marginally stable in the absence of cooling.  
How does introducing finite cooling destabilize the system? 
Inserting Eq. \ref{qcrit_def} into Eq. \ref{inviscid_max} and setting
$\theta=0$, we find 
%In this case the coefficient of 
%$S_*$ in Eq. \ref{inviscid_max} vanish, so that  
\begin{align}\label{sstar}
  S_*^3 = \left(\frac{1}{\gamma Q_\mathrm{crit}^2 \beta}\right). 
\end{align}
The maximum growth rate, $S_*\propto \beta^{-1/3}$, smoothly
increases with decreasing $\beta$. Instability does not require the
cooling time to be below some 
critical value.

@@

\subsection{Viscous disk}\label{2dvisc}
We now consider a viscous disk with 
$\mu=-1,\,\lambda=0$ as discussed in \S\ref{visc_model}. We first
check in Appendix \ref{gammie_check} that our dispersion relation
reduces to that of \cite{gammie96} in the appropriate limit (namely
$\beta\to\infty$).

To see the effect of cooling, we simplify the dispersion relation,
Eq. \ref{thindisk}, by assuming $|\beta S|\ll 1$. Then  
for $|K| \to 0$ we find
\begin{align}\label{gammie_smallk}
%  S \sim \frac{\alpha |K|^3}{(2-q)Q},%derived assuming |beta*S|<<1 
%  S \sim \frac{\alpha K^2\left[2|K|Q^{-1} - \gamma K^2 + 2\alpha q
%    (2-q)(\gamma-1)K^2\right]}{2(2-q)+\gamma K^2 - 2|K|Q^{-1}} %derived
                                %assuming |\beta S|>>1
  S\simeq \frac{\alpha K^2}{2(2-q)}\left(\frac{2|K|}{Q} - \theta
  K^2\right), 
\end{align}
which coincides with \citeauthor{gammie96}'s Eq. 18 for vanishing
wavenumber. For $|K|\to\infty$ we find
\begin{align}\label{gammie_bigk}
  %  S \sim \frac{2}{Q|K|}\left(\frac{4}{3}\alpha + \gamma\beta\right)^{-1}.
  S \simeq\left(\frac{2}{Q|K|} - \theta\right)\left(\frac{4}{3}\alpha + 
  \alpha_b + \gamma\beta\right)^{-1}.
\end{align}

>>>>

\section{Two-dimensional disks with beta cooling}\label{result_2d}
We begin in the 2D limit with beta cooling  to make connection 
with previous studies. The disk
material assumed to be confined to the midplane so $\delta v_z=0$. 
We make the replacements 
$\rho \to \Sigma$ and $P\to\Pi$, the vertically-integrated
pressure, and set $\Gamma=1$ without loss of generality. 
%The perturbations $\delta\Sigma$, 
%$\delta\Pi$, $\delta v_x$ and $\delta v_y$ have no vertical
%dependence and $\delta v_z=0$. 
The gravitational potential perturbation remains 3D and its midplane
value is given by    
\begin{align}
  \delta \Phi(z=0) = -\frac{2 \pi G}{|k|}\delta\Sigma
\end{align}
\citep{shu70}.

@@

\subsection{Inviscid limit}\label{2d_inviscid}
Here we neglect viscosity by setting 
$\alpha = \alpha_b = 0$. This implies an external,
time-independent heat source must be invoked to balance the imposed
cooling rate in equilibrium ($\mathcal{H}_\mathrm{ext}\neq 0$). Alternatively, we are assuming 
viscosity is only important for setting up thermal equilibrium, and
can be neglected for the pertubations. This eliminates viscous GI and 
allows us to quantify the \emph{sole} effect of cooling.

@@

Eq. \ref{inviscid} is a cubic equation in $S$. Applying the
Routh-Hurwitz criterion, we find that stability is ensured if 
\begin{align}
  \gamma > \theta \quad \text{and} \quad 
  Q > \frac{1}{\sqrt{2\theta(2-q)}} 
\end{align}
are both satisfied. Thus, the question of stability does not depend on
the cooling rate $\beta$ (provided it is non-zero and finite). The
system becomes more unstable with decreasing $\theta$.

@@

\subsubsection{$\theta = 0$ and $Q=Q_\mathrm{crit}$}  
Beta cooling with $\theta=0$ is the standard choice for numerical
simulations of self-gravitating disks \citep{gammie01}. There, 
a steady state corresponds to the imposed cooling entirely balanced by heat generated through 
gravito-turbulence. In the present framework, this gravito-turbulence 
is assumed to provide a constant 
background heating ($\mathcal{H}_\mathrm{ext}$) and does
not enter the linearized equations. Cooling, however, is still imposed in the perturbed state. 
For $\theta=0$ the system is unconditionally unstable for
finite $Q$. The most unstable wavelength decreases with decreasing
$\beta$. The $\beta\to0$ limit corresponds to a pressureless disk (not
merely isothermal).

@@

Specifically, let us consider a disk with 
\begin{align}\label{qcrit_def}
  \qtwo = \frac{1}{\sqrt{2\gamma(2-q)}}\equiv Q_{\mathrm{crit}},
\end{align} 
so that it is marginally stable in the absence of cooling.  
How does finite cooling destabilize the disk?  
Inserting Eq. \ref{qcrit_def} into Eq. \ref{inviscid_max} and setting
$\theta=0$, we find 
%In this case the coefficient of 
%$S_*$ in Eq. \ref{inviscid_max} vanish, so that  
\begin{align}\label{sstar}
  S_*^3 = \left(\frac{1}{\gamma Q_\mathrm{crit}^2 \beta}\right). 
\end{align}
The maximum growth rate, $S_*\propto \beta^{-1/3}$, smoothly
increases with decreasing $\beta$. Instability does not require the
cooling time to be below some critical value.

@@

\subsection{Viscous disk}\label{2dvisc}
We now consider a viscous disk with 
$\mu=-1,\,\lambda=0$ as discussed in \S\ref{visc_model}. We first
check in Appendix \ref{gammie_check} that our dispersion relation
reduces previous results in the adiabatic/isothermal limit.

To see the effect of finite cooling, we simplify the dispersion relation,
Eq. \ref{thindisk}, by assuming $|\beta S|\ll 1$. Then  
for $|K| \to 0$ we find
\begin{align}\label{gammie_smallk}
%  S \sim \frac{\alpha |K|^3}{(2-q)Q},%derived assuming |beta*S|<<1 
%  S \sim \frac{\alpha K^2\left[2|K|Q^{-1} - \gamma K^2 + 2\alpha q
%    (2-q)(\gamma-1)K^2\right]}{2(2-q)+\gamma K^2 - 2|K|Q^{-1}} %derived
                                %assuming |\beta S|>>1
  S\simeq \frac{\alpha K^2}{2(2-q)}\left(\frac{2|K|}{Q} - \theta
  K^2\right), 
\end{align}
which coincides with \citeauthor{gammie96}'s Eq. 18 for vanishing
wavenumber. For $|K|\to\infty$ we find
\begin{align}\label{gammie_bigk}
  %  S \sim \frac{2}{Q|K|}\left(\frac{4}{3}\alpha + \gamma\beta\right)^{-1}.
  S \simeq\left(\frac{2}{Q|K|} - \theta\right)\left(\frac{4}{3}\alpha + 
  \alpha_b + \gamma\beta\right)^{-1}.
\end{align}
